Polymethylation scores results

### 2021-10-05 Reports/Results.Rmd
editing results.rmd

```{r temp read n data}
codedir=''
```

@@

# Coding tasks

- X replace epidish cell estimation with ewastools Middleton

- check with jonah about final sample size and exclusions
- determine what to do w/ 1 sample missing clock data (related to above)
- rerun clocks i plan on using on all samples

- make this an entire story line, combined supplemental and non supplemental

@@

# Supplemental Table 1 - Comparison of 450K subsample to Fragile Families cohort

# Supplemental Figure 1 - Cell type proportion estimation

# Supplemental Figure 2 - Choice of beta methylation transformation

# Supplemental Figure 3 - Distribution and correlation of methylation summary measures

#Supplemental Figure 4 - Change in methylation summary measures between visits

@@

fig4a= global_roc%>%
  dplyr::select(predictors, covariates, methylation, childteen, test)%>%
  mutate(modeltype=ifelse(str_detect(predictors, '\\+'), covariates, 'No covariates'))%>%
  unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') &
           covariates=='')%>%
  ggplot(aes(x=spec, y=sens, color=methylation))+geom_line()+
  facet_wrap(childteen~., labeller=labeller(childteen=age_labels), nrow=1)+
  scale_color_discrete('', labels=my_outcomes)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=2, byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()

>>>>

```{r temp read n data}
codedir=''
datadir="/nfs/turbo/bakulski1/People/blostein/FF_methylation/Data"
load(paste0(datadir, '/CreatedData/allPhenoData.Rdata'))
load(paste0(datadir, '/CreatedData/allmethyl.Rdata'))
load(paste0(datadir, '/CreatedData/completeCasemethyl.Rdata'))
```

@@

# Current issues

@@

```{r, eval=T}
#Make Labels
n_fullFF<-paste0("Fragile Families cohort \n", "n=", FF_labeled %>% nrow())
n_methylQC<-paste0("Cohort with 450K methylation data \n", "n=", methylCohort$idnum%>%unique()%>%length(), " individuals with m=", methylCohort$MethID%>%length(), " samples")

n_QCloss<-paste0("Methylation data QC: ", 
                 pdqc_all%>%filter(probe_fail_10==T & childteen!='M')%>%nrow(), " samples / ", 
                 pdqc_all%>%filter(probe_fail_10==T & childteen!='M')%>%dplyr::select(idnum)%>%unique()%>%nrow(), 
                 " individuals with >10% of probes failing dropped, \n",
                 pdqc_all%>%filter(sex_new!=predicted_sex & probe_fail_10==F & childteen!='M')%>%nrow(), " samples / ", 
                 pdqc_all%>%filter(sex_new!=predicted_sex & probe_fail_10==F & childteen!='M')%>%dplyr::select(idnum)%>%unique()%>%nrow(),
                 " individuals with discordant sex samples dropped, \n",
                 pdqc_all%>%filter(predicted_sex_outlier==T & sex_new==predicted_sex & probe_fail_10==F & childteen!='M')%>%nrow(), ' samples / ',
                  pdqc_all%>%filter(predicted_sex_outlier==T & sex_new==predicted_sex & probe_fail_10==F & childteen!='M')%>%dplyr::select(idnum)%>%unique()%>%nrow(), 
                  ' individuals with abnormal sex chromosome intensity dropped & \n',
                 length(reps_remove), " duplicate samples dropped")

n_methylFF<-paste0("Cohort with quality controlled 450K data \n","n=", myFF %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals with m=", myFF%>%nrow(), " samples")

n_baseexclusions<-paste0("n individuals=",  
                         myFF %>% filter_at(all_of(basemodelvar), any_vars((. %in%c("Missing", "Missing PC data"))))%>% dplyr::select(id) %>% unique() %>% nrow(), " \ m samples=", 
                         myFF %>% filter_at(all_of(basemodelvar), any_vars((. %in%c("Missing", "Missing PC data")))) %>% nrow(), " missing principal components of ancestry; \n n individuals=",
                         myFF %>% filter_at(all_of(basemodelvar), any_vars((is.na(.))))%>%dplyr::select(id)%>%unique()%>%nrow(), " individuals \ m=", myFF %>% filter_at(all_of(basemodelvar), any_vars((is.na(.))))%>%nrow(), " samples missing child age at maternal/PCG interview")

n_prenatalexposures<-paste0("Prenatal exposures models: \n","n=", prenatalexdata %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals with m=", prenatalexdata %>%nrow(), " samples")
n_prenatalexclusions<-paste0("n=", basemodeldata %>% filter_at(all_of(prenatalexposurevar), any_vars(. %in% c("Missing"))) %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals & m=", basemodeldata %>% filter_at(all_of(prenatalexposurevar), any_vars(. %in% c("Missing"))) %>% nrow(), " samples missing \n data on maternal prenatal alcohol and drug use")

n_secondhandSmk<-paste0("Secondhand smoke exposure models: \n","n=", secondhandsmkdata %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals with m=", secondhandsmkdata %>%nrow(), " samples")
n_secondhandexclusions<-paste0("n=", prenatalexdata$id[!(prenatalexdata$id %in% secondhandsmkdata$id)] %>% unique()%>% length(), " individuals & m=", prenatalexdata %>% filter_at(all_of(secondhandsmokevars), any_vars(. %in% c("Missing"))) %>% nrow(), " samples missing data \n on maternal or primary care giver postnatal smoking")

n_Child<-paste0('n=', secondhandsmkdata %>%filter(childteen=='C')  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', child_nosmoke %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', child_smoke%>%nrow(), " individual \n smoking at 9 year visit \n or missing focal child smoking data")
n_Teen<-paste0('n=', secondhandsmkdata %>%filter(childteen=='T')   %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', teen_nosmoke %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', teen_smoke%>%nrow(), " individuals \n smoking at 15 year visit \n or missing focal child smoking data")

graph1<-DiagrammeR::grViz("
   digraph graph2 {
   compound=true;
   graph [layout = dot]

   node [shape = rectangle, width=4]
   a [label = '@@1']
   b [label = '@@2']
   c [label = '@@3']
   d [label = '@@4']
   e [label = '@@5']
   f [label = '@@6']
   g [label = '@@7']
   h [label = '@@8']
   i [label = '@@9', group=g1]
   j [label = '@@10']
   k [label = '@@16']
   l [label = '@@17']
   m [label = '@@19']

   #fake nodes for connecting to edges
   node [shape=point, width=0.01, height=0.01]
   a1
   b1
   c1
   d1
#   e1
  f1 [group=g1]
  g1
  k1

  #exclusion nodes
  node [shape = rectangle, width=4]
  aa [label = '@@20']
  bb [label = '@@11']
  cc [label = '@@12']
  dd [label = '@@13']
  ff [label = '@@14', width=3]
  gg [label = '@@15', width=3]
  kk [label = '@@18', width=3]

  #draw connections
  a->m
  m->a1 [arrowhead=none]
  a1->b
  b->b1 [arrowhead=none]
  b1->c
  c->c1 [arrowhead=none]
  c1->d
  d->d1 [arrowhead=none]
  d1->e
  e->f; e->g; e->k
{rank=same; h->e [dir=back, minlen=2, style=dashed]}
  f->f1 [arrowhead=none]
  g->g1 [arrowhead=none]
  k->k1 [arrowhead=none]
  f1->i 
  g1->j 
  k1->l

{rank=same; a1->aa [minlen=2]}   
{rank=same; b1->bb [minlen=2]} 
{rank=same; c1->cc [minlen=2]}
{rank=same; d1->dd [minlen=2]}
{rank=same; f1->ff [minlen=2]}
{rank=same; g1->gg [minlen=2]}
{rank=same; k1->kk [minlen=2]}

#invisible constraints to force ordering
ff->i [style=invis]
   }

   [1]: n_fullFF
   [2]: n_methylFF
   [3]: n_basemodel
   [4]: n_prenatalexposures
   [5]: n_secondhandSmk
   [6]: n_Child
   [7]: n_Teen
   [8]: n_sensFatherSmoking
   [9]: n_ChildSmkexcluded
   [10]: n_TeenSmkexcluded
   [11]: n_baseexclusions
   [12]: n_prenatalexclusions
   [13]: n_secondhandexclusions
   [14]: n_ChildSmkexclusions
   [15]: n_TeenSmkexclusions
   [16]: n_twovisits
   [17]: n_twovisitsSmkExcluded
   [18]: n_twovisitSmkExclusions
   [19]: n_methylQC
   [20]: n_QCloss
")
```

# Supplemental Table 1 - Comparison of excluded sample to included sample

```{r}
myFF$MethylData2<-ifelse(myFF$MethID %in% completecase$MethID, 'Final analysis group', 'Not in final analysis group')
set_label(myFF$MethylData2)<-'Exclusions based on missing data'
myFF=left_join(myFF, methyldata%>%mutate(idnum=as.character(idnum)))
methyl2<-createTable(compareGroups(MethylData2~smkPreg_binary+globalmethylation+pediatric+Epithelial.cells_saliva+Leukocytes_saliva+anynewborn_center+SSnewbornCT_center+SSolder_center+cm1inpov+ancestry+cm1bsex+m1g2_YesNoPreg+m1g3_YesNoPreg+PostnatalMaternalSmokingAny+SmkAtVisitPastmonth, data=myFF), show.all=T)
export2md(strataTable(methyl2, 'childteen'), caption="Exclusion table of subset with quality controlled methylation data vs final analytic subset for main predictor, outcomes and key demographic variables")
```

# Supplemental Figure 2 - Correlation of methylation summary measures

```{r}
library(Hmisc)
library(corrplot)
res2<-completecase%>%dplyr::select(globalmethylation, pediatric, grim, Epithelial.cells_saliva, Leukocytes_saliva, anynewborn_center, SSnewbornCT_center, SSolder_center, cg05575921)%>%as.matrix(.)%>%rcorr()
corrplot.mixed(res2$r, upper="ellipse", lower='number', order='hclust')
```

#Supplemental Figure 3 - Change in methylation summary measures between visits

```{r}

@@

fig4a= global_roc%>%
  dplyr::select(predictors, covariates, methylation, childteen, test)%>%
  mutate(modeltype=ifelse(methylation=='', 'Base model', paste0('Base model+', methylation)))%>%
  unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|SSolder|cg055|global|pediatric'))%>%
  ggplot(aes(x=spec, y=sens, color=modeltype))+geom_line()+
  facet_wrap(childteen~., labeller=labeller(childteen=age_labels), nrow=1)+
  scale_color_discrete('', labels=my_outcomes)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=2, byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()
